Can We Crunch the Social Graph on the Cheap?

### 2011-01-24 doc/hotpar2011/evaluation.tex

To study memory concurrency in commodity systems and begin building a runtime, we did experiments that approximate large irregular applications. The experiments consist of multiple threads each walking a number of separate linked lists. Each memory reference of a node is highly likely to result in an access to main memory, and we control the number of concurrent requests offered by a thread by varying the number of lists each thread walks. \todo{Brandon: somewhere we might explicitly say that this experiment involves no synchronization on list nodes, and relate to plan to build this as mentioned in the components of runtime section} Our experiments were run on a Xeon X5660, which uses the Nehalem microarchitecture, and has 2 sockets, 6 cores per socket, and 12GB of DRAM per socket. For the results listed, memory allocations were pinned to socket0's DRAM. Memory references from socket1's cores must be satisfied by sending requests over the QPI link between the sockets. \todo{Brandon: is it important to mention that for isolation we use huge pages to reduce TLB misses?}
\subsection{Results}

\todo{Might still consider the reqrate v. conc miss graph, because of showing the relation of performance going up as there is more concurrency, or else make sure to describe this clearly in the text}.
\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{figures/multi-listwalk-maxrates-edited.pdf}
	\end{center}
	\caption{The max rates for processing list nodes with varying numbers of cores. Each core walks one list, and each node access results in a reference to main memory. All memory is local to socket0. Cores 1-6 are on socket0, and cores 7-12 are on socket1.}
	\label{fig:listwalk-maxrates}
\end{figure}

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{figures/multi-listwalk-outflats-edited.pdf}
	\end{center}
	\caption{This is my caption about multi-listwak-outflats-edited, it shows the number of misses when rate plateaus...}
	\label{fig:listwalk-outflats}
\end{figure}

\todo{choose best phrase to use for outstanding/concurrent miss/refs/reads and use consistently unless there is a place one makes most sense}
	 Figure~\ref{fig:listwalk-outflats} shows the total number of outstanding misses the system can handle before performance plateaus. We found that for a single core, bandwidth plateaus once there are 10 concurrent misses; this is because the L1 cache line-buffer, which keeps track of outstanding misses of the core, is size 10. If concurrency is limited, there is less latency tolerance and performance is limited.

@@

\begin{figure*}
	\begin{center}
 		\includegraphics[width=0.5\textwidth]{figures/multi-core-green-thread-1gb.pdf}
	\end{center}
	\caption{}
	\label{fig:listwalk-green-threads}
\end{figure*}

@@

	On the Nehalem, the QPI link to the network interface provides a bandwidth 12.8GB/s each way. This as much as 200Mref/s for 64-byte messages. For uniformly random distributed references across the entire shared memory space, we can expect half of this outgoing bandwidth will go towards satisfying remote requests, leaving half for making requests.  So, the cores can put out a bandwidth of 100Mref/s at full utilization of the link. Recent networks have suggested a max issue rate of 30M-90M messages per second is attainable \todo{cite mellanox press release and/or XT3/Eldorado} and this seems to be gradually increasing, so we use 100Mref/s as a good approximation to work with. Assuming a round-trip network latency of $1-2\mu s$, then by Little's Law, the number of outstanding references is $100Mref/s*1-2\mu s = 100-200$. \todo{Redo this calculation with proper numbers latency} This is the amount of memory concurrency that needs to exist for the latency of memory requests to be hidden when there is a rate of 100Mref/s.
We consider two issues: first, can we support 100-200 misses?, and second, can the caches support enough contexts? 
In order to keep the cores busy, there must be sufficient memory concurrency, and the number of contexts required to achieve this will vary. As there are more contexts in a core, there are more workings sets, including both data and coroutine stacks, that should fit in the cache to maintain performance.

	Can we support 100-200 misses? If the remote requests are reads that go to the global queue, then the number of outstanding references is limited to about 36. If the remote requests do not need to be stored in the global queue, then the line-fill buffers of the cores can support a total of $6 cores * 10 ref/core = 60 references$, still not quite enough. If, however, prefetches are messages--communicated with memory stores--requesting a piece of remote memory then many more outstanding references could be supported, either with a software data structure or a memory management device implemented in an FPGA.
	We performed some experiments with green threads on listwalk, where we simulate the network latency by not allowing a coroutine to continue after it does a prefetch-and-switch until some specified amount of time has passed.  As seen in Figure \todo{ref delay plot} it is still possible to achieve the maximum throughput: it just requires more contexts to supply enough concurrent references, as expected. This simulation is of course limited both by the fact that prefetched data is actually sitting in cache longer than it should (increasing the effective working set size) and that the memory system still caps performance, but it is a useful validation of adding software contexts to achieve the level of concurrency that is necessary.

>>>>

To study memory concurrency in commodity systems and begin building a runtime, we did experiments that approximate large irregular applications. The experiments consist of multiple threads each walking a number of separate linked lists. Each memory reference of a node is highly likely to result in an access to main memory, and we control the number of concurrent requests offered by a thread by varying the number of lists each thread walks. \todo{Brandon: somewhere we might explicitly say that this experiment involves no synchronization on list nodes, and relate to plan to build this as mentioned in the components of runtime section} Our experiments were run on a Xeon X5660, which uses the Nehalem microarchitecture, and has 2 sockets, 6 cores per socket, and 12GB of DRAM per socket. For the results listed, memory allocations were pinned to the DRAM of the first socket (``socket0"). Memory references from cores of socket1 must be satisfied by sending requests over the QPI link between the sockets. \todo{Brandon: is it important to mention that for isolation we use huge pages to reduce TLB misses?}
\subsection{Results}

\todo{might also include simple bar graph with max 1 socket, max 2 sockets, to membind=0}

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{figures/multi-listwalk-totalconc-edited.pdf}
	\end{center}
	\caption{Throughput of listwalk versus total number of concurrent references offered by the cores. Data points for running 1 core are shown as circles, and data points for running 6 cores are shown as diamonds. Notice that the throughput for a single core levels out at 10 concurrent references. For 4 to 6 cores, the throughput levels off at around 36 references, which seems to be the most memory concurrency a socket can handle.}
	\label{fig:listwalk-totalconc}
\end{figure}

\todo{choose best phrase to use for outstanding/concurrent miss/refs/reads and use consistently unless there is a place one makes most sense}
	 Figure~\ref{fig:listwalk-totalconc} shows the total number of outstanding misses the system can handle before performance plateaus. We found that for a single core, bandwidth plateaus once there are 10 concurrent misses; this is because the L1 cache line-buffer, which keeps track of outstanding misses of the core, is size 10. If concurrency is limited, there is less latency tolerance and performance is limited.

@@

%\begin{figure*}
%	\begin{center}
%		\includegraphics[width=0.5\textwidth]{figures/multi-core-green-thread-1gb.pdf}
%	\end{center}
%	\caption{}
%	\label{fig:listwalk-green-threads}
%\end{figure*}

@@

	On the Nehalem, the QPI link to the network interface provides a bandwidth 12.8GB/s each way. This as much as 200Mref/s for 64-byte messages. For uniformly random distributed references across the entire shared memory space, we can expect half of this outgoing bandwidth will go towards satisfying remote requests, leaving half for making requests.  So, the cores can put out a bandwidth of 100Mref/s at full utilization of the link. Recent networks have suggested a max issue rate of 30M-90M messages per second is attainable \todo{cite mellanox press release and/or XT3/Eldorado} and this seems to be gradually increasing, so we use 100Mref/s as a good approximation to work with. Assuming a round-trip network latency of 1 to 2$\mu s$, then by Little's Law, the number of outstanding references is $(100Mref/s) * $(2 to 3 $\mu s$) = 200 to 300. This is the amount of memory concurrency that needs to exist for the latency of memory requests to be hidden when there is a rate of 100Mref/s.
We consider two issues: first, can we support 200-300 misses?, and second, can the caches support enough contexts? 
\todo{how many contexts required for that much memory concurrency.  Is it app dependent?} In order to keep the cores busy, there must be sufficient memory concurrency, and the number of contexts required to achieve this will vary. As there are more contexts in a core, there are more workings sets, including both data and coroutine stacks, that should fit in the cache to maintain performance.

	Can we support 200-300 misses? If the remote requests are reads that go to the global queue, then the number of outstanding references is limited to about 36. If the remote requests do not need to be stored in the global queue, then the line-fill buffers of the cores can support a total of $6 cores * 10 ref/core = 60 references$, still not quite enough. If, however, prefetches are messages--communicated with memory stores--requesting a piece of remote memory then many more outstanding references could be supported, either with a software data structure or a memory management device implemented in an FPGA.
	\begin{figure}[h]
	\begin{center}
 		\includegraphics[width=0.5\textwidth]{figures/multi-green-delay7800-edited.pdf}
	\end{center}
	\caption{}
	\label{fig:listwalk-green-delay}
\end{figure}
	We performed some experiments with green threads on listwalk, where we simulate the network latency by not allowing a coroutine to continue after it does a prefetch-and-switch until some specified amount of time has passed.  As seen in Figure \todo{ref delay plot} it is still possible to achieve the maximum throughput: it just requires more contexts to supply enough concurrent references, as expected. This simulation is of course limited both by the fact that prefetched data is actually sitting in cache longer than it should (increasing the effective working set size) and that the memory system still caps performance, but it is a useful validation of adding software contexts to achieve the level of concurrency that is necessary.
